Davos: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2023-09-30 paper/main.tex
updates to section 1 text

At another extreme, virtual machines~\cite{Gold74, AltiEtal05, Rose99}
provide a hardware-level simulation of the desired system.  Virtual
machines are typically isolated, such that installing or running
software on a virtual machine does not impact the user's primary
operating system or computing environment.
Containers~\cite[e.g.,][]{Merk14, KurtEtal17} provide a similar
``isolated'' experience. Although containerized environments do not
specify hardware-level operations, they are typically packaged with a
complete operating system, in addition to a complete copy of Python
and any relevant package dependencies. Virtual
environments~\cite[e.g.,][]{Anac12, Eust19} also provide a computing
environment that is largely separated from the user's main
environment. They incorporate a copy of Python and the target
software's dependencies, but virtual environments do not specify or
reproduce an operating system for the runtime environment. Each of
these systems (virtual machines, containers, and virtual environments)
guarantees (to differing degrees---at the hardware level, operating
system level, and Python environment level, respectively) that the
relevant code will run similarly for different users. However, each of
these systems also relies on additional software that can be complex
or resource-intensive to install and use, creating potential barriers
to both contributing to and taking advantage of open science
resources.

We designed Davos to occupy a ``sweet spot'' between these extremes.
Davos is a notebook-installable package that adds functionality to the
default notebook experience. Like standard Jupyter notebooks,
Davos-enhanced notebooks allow researchers to include text, executable
code, and media within a single file. No further setup or installation is
required from the user, beyond what is needed to run standard Jupyter
notebooks. And like virtual environments, Davos provides a convenient
mechanism for fully specifying (and installing, as needed) a complete set of
Python dependencies, including specific package versions, which are contained
and isolated from the rest of the user's system.

>>>>

At another extreme, virtual machines~\cite{Gold74, AltiEtal05, Rose99}
provide a hardware-level simulation of the desired system.  Virtual
machines are typically isolated, such that installing or running
software on a virtual machine does not impact the user's primary
operating system or computing environment.
Containers~\cite[e.g.,][]{Merk14, KurtEtal17} provide a similar
``isolated'' experience. Although containerized environments do not
specify hardware-level operations, they are typically packaged with a
complete operating system, in addition to a complete copy of Python
and any relevant package dependencies. Shareable Python
environments~\cite[e.g.,][]{Anac12} also provide a computing
environment that is largely separated from the user's main
environment. They incorporate a copy of Python and the target
software's dependencies, but do not specify or
reproduce an operating system for the runtime environment.
Virtual environments~\cite[e.g.,][]{BickEtal07, Eust19} work similarly, but
reuse an existing copy of Python rather than bundling their own.
Each of these systems (virtual machines, containers, Python environments, and virtual environments)
guarantees (to differing degrees---at the hardware level, operating
system level, Python environment level, and package environment level, respectively) that the
relevant code will run similarly for different users. However, each of
these systems also relies on additional software that can be complex
or resource-intensive to install and use, creating potential barriers
to both contributing to and taking advantage of open science
resources.

We designed Davos to occupy a ``sweet spot'' within this spectrum.
Davos is a notebook-installable package that adds functionality to the
default notebook experience. Like standard Jupyter notebooks,
Davos-enhanced notebooks allow researchers to include text, executable
code, and media within a single file. No further setup or installation is
required from the user, beyond what is needed to run a standard Jupyter
notebook. And like virtual environments, Davos provides a convenient
mechanism for fully specifying (and installing, as needed) a complete set of
Python dependencies, including specific package versions, which are contained
and isolated from the rest of the user's system.
